Efficient pedigree recording for fast population genetics simulation

### 2017-12-02 forwards_paper.tex
Mention forwsim and anafits

However, it is thought that most genetic variation is selectively neutral (or nearly so).
By definition, the alleles carried by individuals in a population at neutral sites
do not affect the population process.
For this reason, if one records the entire genealogical history of a population over the course of a simulation,
simply laying down neutral mutations on top of that history afterwards
is equivalent to having generated them during the simulation.
Precisely, we would need to know the genealogical tree relating all sampled individuals
at each position along the genome.
In this paper, we describe how algorithmic tools and data structures developed for the
coalescent simulator \msprime{}
can be used to efficiently record, and later process, this history.
In so doing we record the \emph{population pedigree} --
the complete history of parent-offspring relationships of an entire population
going back to a remote time including the genetic outcomes of each ancestral
meiosis.
This embellished pedigree contains all the information necessary
to construct the genealogical tree that relates each individual to each other
at each position on the genome, i.e., the \emph{tree
sequence}~\citep{kelleher2016efficient}.
Combined with ancestral genotypes and the origins of new mutations,
it also completely specifies the genomic sequence of any individual in the population at any time.
This is much more than we need to know, however,
so we discard all information irrelevant to the genetic history
of the \emph{sampled} individuals,
which results in considerable savings.
Another way of representing this same information
is known as the \emph{ancestral recombination graph} ---
the {ARG} \citep{griffiths1997ancestral},
where nodes are common ancestor or recombination events ---
which has been the subject of substantial study
under the assumptions of coalescent
theory~\citep{wiuf1997number,wiuf1999ancestry,marjoram2006coalescent,wilton2015smc}.

>>>>

However, it is thought that most genetic variation is selectively neutral (or nearly so).
By definition, the alleles carried by individuals in a population at neutral sites
do not affect the population process.
For this reason, if one records the entire genealogical history of a population over the course of a simulation,
simply laying down neutral mutations on top of that history afterwards
is equivalent to having generated them during the simulation.
Precisely, we would need to know the genealogical tree relating all sampled individuals
at each position along the genome.
This basic insight has already led development of specialized forward simulators
(e.g., ~\citet{forwsim}, ``Ancestry aware forward in time simulator" (anafits))
that track only recombination and neutral haplotypes.

In this paper, we describe how algorithmic tools and data structures developed for the
coalescent simulator \msprime{}
can be used to efficiently record, and later process, this history with
\emph{any} forward-time simulator.
In so doing we record the \emph{population pedigree} --
the complete history of parent-offspring relationships of an entire population
going back to a remote time including the genetic outcomes of each ancestral
meiosis.
This embellished pedigree contains all the information necessary
to construct the genealogical tree that relates each individual to each other
at each position on the genome, i.e., the \emph{tree
sequence}~\citep{kelleher2016efficient}.
Combined with ancestral genotypes and the origins of new mutations,
it also completely specifies the genomic sequence of any individual in the population at any time.
This is much more than we need to know, however,
so we discard all information irrelevant to the genetic history
of the \emph{sampled} individuals,
which results in considerable savings.
Another way of representing this same information
is known as the \emph{ancestral recombination graph} ---
the {ARG} \citep{griffiths1997ancestral},
where nodes are common ancestor or recombination events ---
which has been the subject of substantial study
under the assumptions of coalescent
theory~\citep{wiuf1997number,wiuf1999ancestry,marjoram2006coalescent,wilton2015smc}.
